An R Markdown document converted from "02_ValClus.ipynb

### 2025-11-12 scripts/14_Rebuttal.Rmd

---
title: An R Markdown document converted from "17_Rebuttal.ipynb"
output: html_document
---

# Reviewer's comments

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr)
library(purrr)
library(ggplot2)
```

```{r}
options(repr.plot.res = 300)
```

## Reviewer #1

1. The authors identified the five BMI-discordant profiles (DHT, DAL, DLT, DIS, and DHG). However, as they could easily assign these profile names, all these profiles were just characterized by a single biomarker (or a few biomarkers which are well-known to be highly associated with each other, such as TG, LDL, and HDL for the DAL profile) (lines 144–152 of pages 10–11; Figure 1 right panel). Therefore, it is unclear about the advantage of using the identified profiles compared to simply using each biomarker-based classification (with BMI). Based on the authors' aim, I would have expected profiles that were characterized by multiple biomarkers (with various weights) whose associations are not so obvious (e.g., TG vs. CRP). Hence, I have concerns about the conceptual advances of the identified discordant profiles. Even if the authors keep claiming the usefulness of the current discordant profiles, they must present the comparisons between the identified discordant profiles and the single biomarker-based profiles in their downstream analyses (e.g., the additional net benefits by including the profile information), because the current analyses just evaluate the comparisons between with vs. without the discordant profile information.

R/

The reviewer accurately points out that the clusters we have identified are largely characterised by a single biomarker being at values that are higher than the expected for their BMI.

```{r}
load("../data/validclusmod.RData")
```

```{r}
print(validclusmod)
```

```{r}
validclusmod <- validclusmod %>%
    mutate(
        umapmod = map(
            sex,
            ~uwot::load_uwot(paste0("../data/ukb/umapmod_", .x))
        )
    )
print(validclusmod)
```

Let's see what happens when we take an individual that is 1) predominantly concordant, and then 2) increase triglycerides only, or 3) increase LDL and decrease HDL (what happens in DAL):

```{r}
NewInd <- tibble(
    eid = 1:3,
    whr = 1,
    sbp = 140, 
    dbp = 80, 
    alt = 28, 
    scr = 70, 
    crp = 2.5, 
    hdl = c(1.3, 1.3, 1), 
    tg = c(1.8, 3.1, 3.1),
    ldl = c(3.5, 3.5, 4.5),
    fg = 5
)
validclusmod %>%
    filter(sex == "Male") %>%
    transmute(
        resids = map(
            residmod,
            ~.x %>%
                transmute(
                    Biomarker,
                    PredValue = Intercept + (bmi * 30) + (age * 55),
                    ResidValue = pmap(
                        list(Biomarker, PredValue, SDRes),
                        function(BM, PV, SDV, NewD = NewInd){
                            (NewD[,BM,drop=TRUE] - PV) / SDV
                        }
                    )
                ) %>%
                select(Biomarker, ResidValue) %>%
                pivot_wider(names_from = Biomarker, values_from = ResidValue) %>%
                unnest(everything()) %>%
                as.matrix
        ),
        probs = map2(
            resids, clusmod,
            function(RESIDDF, CLUSMOD){
                BIOMARKERS <- c("whr", "sbp", "dbp", "alt", "scr", "crp", "hdl", "tg", "ldl", "fg")
                RESIDDF <- RESIDDF[,BIOMARKERS]
                mus <- CLUSMOD$validclus_centers
                mus <- map(mus, function(mu){ mu[BIOMARKERS] })
                covmats <- CLUSMOD$validclus_covmats
                covmats <- map(covmats, function(covmat){ covmat[BIOMARKERS, BIOMARKERS] })
                wts <- CLUSMOD$validclus_weights
                pdfs <- mapply(
                    function(mu, covmat){ mvtnorm::dmvnorm(RESIDDF, mu, covmat) },
                    mus, covmats
                )
                L <- pdfs %*% diag(wts)
                colnames(L) <- paste0("prob", CLUSMOD$validclus_name)
                tibble(data.frame(L / rowSums(L)))
            }
        )
    ) %>%
    pull(probs) %>%
    pluck(1) %>%
    round(2)
```

Now let's take the DHG profile and show what happens when a 1) concordant person has 2) isolated hyperglycaemia or 2) has also lower LDL and HDL - the other caracteristic of DHG.

```{r}
NewInd <- tibble(
    eid = 1:3,
    whr = 0.99,
    sbp = 140, 
    dbp = 80, 
    alt = 28, 
    scr = 70, 
    crp = 2.1, 
    hdl = c(1.3, 1.3, 1),
    tg = 1.8,
    ldl = c(3.5, 3.5, 2.5),
    fg = c(4.9, 6.7, 6.7)
)
validclusmod %>%
    filter(sex == "Male") %>%
    transmute(
        resids = map(
            residmod,
            ~.x %>%
                transmute(
                    Biomarker,
                    PredValue = Intercept + (bmi * 30) + (age * 55),
                    ResidValue = pmap(
                        list(Biomarker, PredValue, SDRes),
                        function(BM, PV, SDV, NewD = NewInd){
                            (NewD[,BM,drop=TRUE] - PV) / SDV
                        }
                    )
                ) %>%
                select(Biomarker, ResidValue) %>%
                pivot_wider(names_from = Biomarker, values_from = ResidValue) %>%
                unnest(everything()) %>%
                as.matrix
        ),
        probs = map2(
            resids, clusmod,
            function(RESIDDF, CLUSMOD){
                BIOMARKERS <- c("whr", "sbp", "dbp", "alt", "scr", "crp", "hdl", "tg", "ldl", "fg")
                RESIDDF <- RESIDDF[,BIOMARKERS]
                mus <- CLUSMOD$validclus_centers
                mus <- map(mus, function(mu){ mu[BIOMARKERS] })
                covmats <- CLUSMOD$validclus_covmats
                covmats <- map(covmats, function(covmat){ covmat[BIOMARKERS, BIOMARKERS] })
                wts <- CLUSMOD$validclus_weights
                pdfs <- mapply(
                    function(mu, covmat){ mvtnorm::dmvnorm(RESIDDF, mu, covmat) },
                    mus, covmats
                )
                L <- pdfs %*% diag(wts)
                colnames(L) <- paste0("prob", CLUSMOD$validclus_name)
                tibble(data.frame(L / rowSums(L)))
            }
        )
    ) %>%
    pull(probs) %>%
    pluck(1) %>%
    round(2)
```

Additional information of discordant profiles to models containing single biomarker discordances - represented by biomarker-BMI ratios.

```{r}
load("../data/ukb/strat_dat.RData")
```

```{r}
map(strat_dat, head)
```

Then the covariate data:

```{r}
covar_dat <- read_tsv("../data/covar_dat.tsv", show_col_types = FALSE)
head(covar_dat)
```

Then the survival data:

```{r}
survmacedat <- read_tsv("../data/survmacedat.tsv", show_col_types = FALSE)
head(survmacedat)
```

```{r}
survdmdat <- read_tsv("../data/survdmdat.tsv", show_col_types = FALSE)
head(survdmdat)
```

Calculate cluster probabilities:

```{r}
BIOMARKERS <- c("whr", "sbp", "dbp", "alt", "scr", "crp", "hdl", "tg", "ldl", "fg")
BODYSIZEINDEX <- "bmi"
COVARIATES <- c("age", "smoking")
DISEASES <- c("HT", "CHD", "Stroke", "PAD", "CKD", "LiverFailure", "RA", "T2D", "T1D")
MEDICATION <- c("Insulin", "AntiDM", "AntiHT", "LipidLower")
```

```{r}
clusterdfs <- validclusmod %>%
    mutate(
        Data = map(sex, ~strat_dat[[.x]]),
        resids = map2(
            Data, residmod,
            function(DATA, RESIDMOD){
                sapply(
                    BIOMARKERS,
                    function(biomarker){
                        PREDICTDAT <- as.matrix(cbind(Intercept = 1, DATA[,c(BODYSIZEINDEX, COVARIATES)]))
                        MODELDAT <- filter(RESIDMOD, Biomarker == biomarker)
                        RESIDSD <- MODELDAT$SDRes
                        COEFDAT <- MODELDAT[,c("Intercept", BODYSIZEINDEX, COVARIATES)]
                        COEFDAT <- t(COEFDAT)
                        PREDVAL <- as.vector(PREDICTDAT %*% COEFDAT)
                        RESIDVAL <- DATA[[biomarker]] - PREDVAL
                        RESIDVAL / RESIDSD
                    }
                )
            }
        ),
        probs = map2(
            resids, clusmod,
            function(RESIDDF, CLUSMOD){
                mus <- CLUSMOD$validclus_centers
                mus <- map(mus, function(mu){ mu[BIOMARKERS] })
                covmats <- CLUSMOD$validclus_covmats
                covmats <- map(covmats, function(covmat){ covmat[BIOMARKERS, BIOMARKERS] })
                wts <- CLUSMOD$validclus_weights
                pdfs <- mapply(function(mu, covmat){ mvtnorm::dmvnorm(RESIDDF, mu, covmat) }, mus, covmats)
                L <- pdfs %*% diag(wts)
                colnames(L) <- paste0("prob", CLUSMOD$validclus_name)
                data.frame(L / rowSums(L))
            }
        ),
        Data = map2(Data, probs, bind_cols),
        resids = NULL, probs = NULL,
        Data = map(Data, inner_join, covar_dat, by = "eid")
    )
print(clusterdfs)
```

Models for MACE:

```{r}
clusmacedfs <- clusterdfs %>%
    mutate(
        Data = map(Data, inner_join, survmacedat, by = "eid"),
        Data = map(Data, filter, CHD == 0, Stroke == 0, PAD == 0),
        Data = map2(
            Data, sex,
            function(DATA, SEX){
                MODELDAT <- transmute(
                    DATA,
                    outcome_value,
                    outcome_timeyrs,
                    age,
                    smoking,
                    bmi,
                    sbp,
                    tchol = hdl + ldl + (tg/2.2),
                    hdl,
                    age_smoking = age*smoking,
                    age_sbp = age*sbp,
                    age_tchol = age*tchol,
                    age_hdl = age*hdl,
                    fg,
                    scr = scr / 88.42,
                    alpha = case_when(SEX == "Female" ~ -0.241, SEX == "Male" ~ -0.302),
                    kappa = case_when(SEX == "Female" ~ 0.7, SEX == "Male" ~ 0.9),
                    creat_kappa = scr / kappa,
                    minkappa = pmin(creat_kappa, 1), 
                    maxkappa = pmax(creat_kappa, 1),
                    minkappa_alpha = minkappa^alpha,
                    maxkappa_exp = maxkappa^(-1.200),
                    age_term = 0.9938^age,
                    sex_term = case_when(SEX == "Female" ~ 1.012, SEX == "Male" ~ 1), 
                    egfr = 142 * minkappa_alpha * maxkappa_exp * age_term * sex_term,
                    lnegfr = log(egfr + 0.01),
                    lnegfrsq = lnegfr^2,
                    age_fg = age*fg,
                    age_lnegfr = age*lnegfr,
                    whr, alt, crp,
                    across(
                        any_of(
                            c("T2D", "T2Dage", "HT", "CKD", "LiverFailure", "RA", "T1D",
                              "Insulin", "AntiDM", "AntiHT", "LipidLower")
                        )
                    ),
                    across(all_of(BIOMARKERS), \(x){ x/bmi }, .names = "{.col}_ratio"),
                    across(starts_with("prob"), \(x) ifelse(x < 1e-15, 1e-15, x)),
                    across(starts_with("prob"), \(x) ifelse(x > (1 - 1e-15), (1 - 1e-15), x)),
                    across(starts_with("prob"), \(P) log(P/probBC), .names = "alr{.col}"),
                    alrprobBC = NULL,
                    across(starts_with("prob"), \(P) NULL)
                )
                if(any(names(MODELDAT) == "T2D")){
                    MODELDAT$age_t2d <- MODELDAT$age * MODELDAT$T2D
                }
                select(
                    MODELDAT, 
                    -c(scr, alpha, kappa, creat_kappa, minkappa, maxkappa, 
                       minkappa_alpha, maxkappa_exp, age_term, sex_term, egfr)
                )
            }
        )
    )
print(clusmacedfs)
```

```{r}
clusmacedfs$Data[[1]] %>% names
```

```{r}
clusmacedfs %>%
    mutate(
        NullMod = map(
            Data, 
            ~survival::coxph(
                survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ 1, 
                data = .x
            )
        ),
        mod_base = map(
            Data,
            ~{
                MODELDAT <- select(.x, -starts_with("alrprob"))
                survival::coxph(
                    survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ ., 
                    data = MODELDAT
                )
            }
        ),
        mod_clus = map(
            Data,
            ~survival::coxph(
                survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ ., 
                data = .x
            )
        )
    ) %>%
    transmute(
        sex,
        LL0 = purrr::map_dbl(NullMod, logLik),
        LLBase = purrr::map_dbl(mod_base, logLik),
        NVBase = purrr::map_dbl(mod_base, \(MOD) sum(!is.na(MOD$coefficients))),
        LLBaseCl = purrr::map_dbl(mod_clus, logLik),
        NVBaseCl = purrr::map_dbl(mod_clus, \(MOD) sum(!is.na(MOD$coefficients))),
        LRTstat = 2 * abs(LLBaseCl - LLBase),
        LRTdf = NVBaseCl - NVBase,
        LRTp = stats::pchisq(
            q = LRTstat,
            df = LRTdf,
            lower.tail = FALSE
        ),
        AdeqInd = (LLBase - LL0) / (LLBaseCl - LL0),
        RelExpVar = round(100 * (1 - AdeqInd), 2)
    )
```

For diabetes:

```{r}
clusdmdfs <- clusterdfs %>%
    mutate(
        Data = map(Data, inner_join, survdmdat, by = "eid"),
        Data = map(Data, filter, T2D == 0, T1D == 0, Insulin == 0, AntiDM == 0),
        Data = purrr::map(
            Data,
            function(DATA){
                dplyr::transmute(
                    DATA,
                    outcome_timeyrs, outcome_value,
                    across(
                        any_of(
                            c(
                                BIOMARKERS,
                                BODYSIZEINDEX,
                                COVARIATES,
                                DISEASES[!DISEASES %in% c("T1D", "T2D")],
                                MEDICATION[!MEDICATION %in% c("Insulin", "AntiDM")]
                            )
                        )
                    ),
                    across(all_of(BIOMARKERS), \(x){ x/bmi }, .names = "{.col}_ratio"),
                    across(starts_with("prob"), \(x) ifelse(x < 1e-15, 1e-15, x)),
                    across(starts_with("prob"), \(x) ifelse(x > (1 - 1e-15), (1 - 1e-15), x)),
                    across(starts_with("prob"), \(P) log(P/probBC), .names = "alr{.col}"),
                    alrprobBC = NULL,
                    across(starts_with("prob"), \(P) NULL)
                )
            }
        )
    )
print(clusdmdfs)
```

```{r}
clusdmdfs$Data[[1]] %>% names
```

```{r}
clusdmdfs %>%
    mutate(
        NullMod = map(
            Data, 
            ~survival::coxph(
                survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ 1, 
                data = .x
            )
        ),
        mod_base = map(
            Data,
            ~{
                MODELDAT <- select(.x, -starts_with("alrprob"))
                survival::coxph(
                    survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ ., 
                    data = MODELDAT
                )
            }
        ),
        mod_clus = map(
            Data,
            ~survival::coxph(
                survival::Surv(time = outcome_timeyrs, event = outcome_value) ~ ., 
                data = .x
            )
        )
    ) %>%
    transmute(
        sex,
        LL0 = purrr::map_dbl(NullMod, logLik),
        LLBase = purrr::map_dbl(mod_base, logLik),
        NVBase = purrr::map_dbl(mod_base, \(MOD) sum(!is.na(MOD$coefficients))),
        LLBaseCl = purrr::map_dbl(mod_clus, logLik),
        NVBaseCl = purrr::map_dbl(mod_clus, \(MOD) sum(!is.na(MOD$coefficients))),
        LRTstat = 2 * abs(LLBaseCl - LLBase),
        LRTdf = NVBaseCl - NVBase,
        LRTp = stats::pchisq(
            q = LRTstat,
            df = LRTdf,
            lower.tail = FALSE
        ),
        AdeqInd = (LLBase - LL0) / (LLBaseCl - LL0),
        RelExpVar = round(100 * (1 - AdeqInd), 2)
    )
```

# Projection using other methods

```{r}
load("../data/ukb/residtab.RData")
```

```{r}
map(residtab, head)
```

```{r}
probdfs <- residtab %>%
    imap(
        ~{
            RESIDDF <- residtab[[.y]][,-1]
            CLUSMOD <- validclusmod$clusmod[validclusmod$sex == .y][[1]]
            pdfs <- mapply(
                function(mu, covmat){ mvtnorm::dmvnorm(RESIDDF, mu, covmat) },
                CLUSMOD$validclus_centers, 
                CLUSMOD$validclus_covmats
            )
            L <- pdfs %*% diag(CLUSMOD$validclus_weights)
            colnames(L) <- paste0("prob", CLUSMOD$validclus_name)
            tibble(data.frame(L / rowSums(L)))
        }
    )
```

```{r}
map(probdfs, head)
```

```{r}
maxprobdat <- probdfs %>%
    map(
        ~{
            MAXCLUS <- apply(.x, 1, \(x) colnames(.x)[which.max(x)])
            MAXCLUSP <- apply(.x, 1, max)
            tibble(data.frame(Cluster = MAXCLUS, Prob = MAXCLUSP))
        }
    )
map(maxprobdat, head)
```

```{r}
tsne <- map(
    residtab,
    ~Rtsne::Rtsne(
        .x[,-1], 
        dims = 2,
        perplexity = 30,
        max_iter = 500,
        check_duplicates = FALSE
    )
)
```

### 2025-11-12 scripts/02_ValClus.Rmd
Initial clean commit

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
percsmoke %>%
    filter(currentsmoke == "Yes") %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    ggplot(aes(cohort, Perc)) +
    geom_col(aes(fill = cohort), show.legend = FALSE) +
    scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
    facet_wrap(
        ~sex
    ) +
    labs(x = NULL, y = "%") + 
    theme_bw()
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
numgendesc %>%
    mutate(
        Variable = toupper(Variable),
        trait = Variable,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    group_by(Variable) %>%
    nest %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort)) +
                geom_boxplot(aes(fill = cohort, ymin = p2.5, lower = p25, middle = Median, upper = p75, ymax = p97.5),
                             stat = "identity") +
                scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top", legend.title = element_blank())
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
regcoefs %>%
    filter(term == "bmi") %>%
    mutate(
        Biomarker = toupper(Biomarker),
        trait = Biomarker,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    nest(data = -Biomarker) %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort, Estimate)) +
                geom_hline(yintercept = 0, lty = "dashed") +
                geom_linerange(aes(ymin = lowerCI, ymax = upperCI)) +
                geom_point(aes(color = cohort)) +
                scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top")
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 6)
umap_projs %>%
    mutate(
        cohort = recode_factor(
            factor(cohort), 
            "ukb" = "UKB",
            "rotterdam" = "RS",
            "maastricht" = "MS",
            "ghs" = "GHS"
        )
    ) %>%
    ggplot(aes(X1, X2)) +
    geom_point(aes(color = cohort), shape = ".", show.legend = FALSE) +
    scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = 1) +
    ggh4x::facet_nested_wrap(~cohort + sex, scales = "free", nrow = 2) +
    theme_bw() +
    theme(
        axis.text = element_blank(),
        axis.ticks = element_blank()
    ) +
    labs(x = "UMAP1", y = "UMAP2")
```

@@

```{r}
options(repr.plot.width = 12, repr.plot.height = 15)
result_tab %>%
    transmute(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        ),
        data = map(data, "Cluster_results"),
        sex = map(data, names)
    ) %>%
    unnest(c(sex, data)) %>%
    mutate(
        data = map(data, "clusters"),
        cluster = map(data, names)
    ) %>%
    unnest(cluster) %>%
    mutate(
        data = map2(data, cluster, ~.x[[.y]]),
        center = map(data, "center"),
        trait = map(center, names)
    ) %>%
    select(-data) %>%
    unnest(c(center, trait)) %>%
    mutate(trait = toupper(trait)) %>%
    ggplot(aes(center, trait)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_segment(aes(xend = 0, yend = trait)) +
    geom_point() +
    ggh4x::facet_nested(sex + cohort ~ cluster) +
    labs(x = "Mean residual value (SD units)", y = NULL) +
    theme_bw()
```

@@

```{r}
options(repr.plot.width = 7, repr.plot.height = 4)
validatedtab %>%
    transmute(
        sex,
        from = gsub("cluster_", "U:", cluster_ref),
        to = case_match(
            cohort_test,
            "maastricht" ~ gsub("cluster_", "M:", cluster_test),
            "rotterdam" ~ gsub("cluster_", "R:", cluster_test),
            "ghs" ~ gsub("cluster_", "G:", cluster_test)
        )
    ) %>%
    group_by(sex) %>%
    nest %>%
    mutate(
        net = map(data, ~igraph::graph_from_data_frame(.x, directed = FALSE)),
        netplot = map2(
            net, sex,
            ~.x %>%
                ggraph::ggraph(layout = "igraph", algorithm = "nicely") +
                ggraph::geom_edge_link() +
                ggraph::geom_node_point(aes(fill = gsub("^([A-Z]):.+$", "\\1", name)), 
                                        shape = 21, size = 7) +
                scale_fill_viridis_d(
                    option = "A", begin = .25, end = .75, direction = -1,
                    labels = c(U = "UKB", M = "MS", R = "RS", G = "GHS"),
                    guide = guide_legend(override.aes = list(size = 2))
                ) +
                ggraph::geom_node_text(aes(label = name), color = "white", fontface = "bold", size = 2) +
                labs(title = .y, fill = "Cohort")
        )
    ) %>%
    pull(netplot) %>%
    patchwork::wrap_plots(nrow = 1, guides = "collect") &
    theme(legend.position = "bottom")
```

>>>>

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
percsmoke %>%
    filter(currentsmoke == "Yes") %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    ggplot(aes(cohort, Perc)) +
    geom_col(aes(fill = cohort), show.legend = FALSE) +
    scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
    facet_wrap(
        ~sex
    ) +
    labs(x = NULL, y = "%") + 
    theme_bw()
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
numgendesc %>%
    mutate(
        Variable = toupper(Variable),
        trait = Variable,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    group_by(Variable) %>%
    nest %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort)) +
                geom_boxplot(aes(fill = cohort, ymin = p2.5, lower = p25, middle = Median, upper = p75, ymax = p97.5),
                             stat = "identity") +
                scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top", legend.title = element_blank())
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
regcoefs %>%
    filter(term == "bmi") %>%
    mutate(
        Biomarker = toupper(Biomarker),
        trait = Biomarker,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    nest(data = -Biomarker) %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort, Estimate)) +
                geom_hline(yintercept = 0, lty = "dashed") +
                geom_linerange(aes(ymin = lowerCI, ymax = upperCI)) +
                geom_point(aes(color = cohort)) +
                scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top")
```

@@

```{r}
options(repr.plot.width = 10, repr.plot.height = 6)
umap_projs %>%
    mutate(
        cohort = recode_factor(
            factor(cohort), 
            "ukb" = "UKB",
            "rotterdam" = "RS",
            "maastricht" = "TMS",
            "ghs" = "GHS"
        )
    ) %>%
    ggplot(aes(X1, X2)) +
    geom_point(aes(color = cohort), shape = ".", show.legend = FALSE) +
    scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = 1) +
    ggh4x::facet_nested_wrap(~cohort + sex, scales = "free", nrow = 2) +
    theme_bw() +
    theme(
        axis.text = element_blank(),
        axis.ticks = element_blank()
    ) +
    labs(x = "UMAP1", y = "UMAP2")
```

@@

```{r}
options(repr.plot.width = 12, repr.plot.height = 15)
result_tab %>%
    transmute(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        ),
        data = map(data, "Cluster_results"),
        sex = map(data, names)
    ) %>%
    unnest(c(sex, data)) %>%
    mutate(
        data = map(data, "clusters"),
        cluster = map(data, names)
    ) %>%
    unnest(cluster) %>%
    mutate(
        data = map2(data, cluster, ~.x[[.y]]),
        center = map(data, "center"),
        trait = map(center, names)
    ) %>%
    select(-data) %>%
    unnest(c(center, trait)) %>%
    mutate(trait = toupper(trait)) %>%
    ggplot(aes(center, trait)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_segment(aes(xend = 0, yend = trait)) +
    geom_point() +
    ggh4x::facet_nested(sex + cohort ~ cluster) +
    labs(x = "Mean residual value (SD units)", y = NULL) +
    theme_bw()
```

@@

```{r}
options(repr.plot.width = 7, repr.plot.height = 4)
validatedtab %>%
    transmute(
        sex,
        from = gsub("cluster_", "U:", cluster_ref),
        to = case_match(
            cohort_test,
            "maastricht" ~ gsub("cluster_", "M:", cluster_test),
            "rotterdam" ~ gsub("cluster_", "R:", cluster_test),
            "ghs" ~ gsub("cluster_", "G:", cluster_test)
        )
    ) %>%
    group_by(sex) %>%
    nest %>%
    mutate(
        net = map(data, ~igraph::graph_from_data_frame(.x, directed = FALSE)),
        netplot = map2(
            net, sex,
            ~.x %>%
                ggraph::ggraph(layout = "igraph", algorithm = "nicely") +
                ggraph::geom_edge_link() +
                ggraph::geom_node_point(aes(fill = gsub("^([A-Z]):.+$", "\\1", name)), 
                                        shape = 21, size = 7) +
                scale_fill_viridis_d(
                    option = "A", begin = .25, end = .75, direction = -1,
                    labels = c(U = "UKB", M = "TMS", R = "RS", G = "GHS"),
                    guide = guide_legend(override.aes = list(size = 2))
                ) +
                ggraph::geom_node_text(aes(label = name), color = "white", fontface = "bold", size = 2) +
                labs(title = .y, fill = "Cohort")
        )
    ) %>%
    pull(netplot) %>%
    patchwork::wrap_plots(nrow = 1, guides = "collect") &
    theme(legend.position = "bottom")
```
